A molecular dynamics study of the solvation of \ce{CO_2}, water, and some organic compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-08-03 ionic_liquids.tex
add pressure figure

The initial configurations were obtained with PLAYMOL \cite{playmol}, a software integrated with PACKMOL \cite{Mart_nez_2009}, using the tool \textit{build} to specify geometric data (bond and angle lengths, and dihedral angles), from which PLAYMOL automatically determines the Cartesian coordinates.
We first carried out NVT MD simulations of the isolated ion pairs in order to obtain a minimum energy conformation.
For this, we used the Nos\'{e}-Hoover thermostat \cite{Martyna1992} and a cubic box of length L = 50 {\AA} with non-periodic boundary conditions.
After a short run with $\Delta t$ = $0.5~/text{fs}$, we carried out a ``quenching'' simulation with $\Delta t$ = $1 ~/text{fs}$, in which the temperature is gradually reduced until reaching a value close to $0~/text{K}$.
The resulting coordinates were then specified in PLAYMOL \cite{playmol} to build the configurations of the liquid phases.

@@

Using the software TRAVIS \cite{Brehm_2011}, we obtained the combined distribution functions (CDFs) of the dihedral angles of C1-N2-C5-C6 and C2-N2-C5-C6, as sampled in NPT-MD simulations of 250 ion pairs at $T = 298.0~\text{K}$ and $P = 1.0~\text{bar}$.
In Fig.~\ref{fig:dihedrals_emim} we show the CDFs for each force field of \ce{[emim][B(CN)_4]}.
A first inspection of this figure reveals a similar behavior of the models due to Koller \textit{et al.} \cite{Koller_2012} and Liu \textit{et al.} \cite{Liu_2014}.
This was expected as both are based on the AMBER framework \cite{Cornell_1995}.
In addition, Figs.~\ref{fig:dihedrals_emim}(a) and \ref{fig:dihedrals_emim}(b) are symmetric, which means that atoms C1, N2, and C2 are coplanar.
In Fig.~\ref{fig:dihedrals_emim}(a), the most likely angles can be found in a narrow region around 90$^{\circ}$ and 270$^{\circ}$.
Since these angles correspond to an equivalent configuration in the \textit{united-atom} \ce{[emim]^+}, in the case of Koller \textit{et al.} \cite{Koller_2012} we built all the rigid bodies by specifying 90$^{\circ}$ for the dihedral angle of C6-C5-N2-C1.
The other models show a larger flexibility with high probability regions that extend in a broad range of angle values.
Despite this, it was possible to build more than one type of rigid body by specifying the most probables angles of C6-C5-N2-C1.
In the case of Liu \textit{et al.} \cite{Liu_2014}, these angles correspond to 90$^{\circ}$ and -90$^{\circ}$, while for the model of Batista \textit{et al.} \cite{Batista_2015}, we employed 102.6$^{\circ}$ and -102.6$^{\circ}$.
These two types of rigid bodies were built in equal proportion.
In Fig.~\ref{fig:dihedrals_emim}(d), which corresponds to Weber and Kirchner \cite{Weber_2016}, one sees that the angle As this is the most flexible model, we defined four types of rigid bodies: 150$^{\circ}$, 180$^{\circ}$ and 210$^{\circ}$ were employed to build 30~\%, 25~\% and 30~\% of them, respectively; and 0$^{\circ}$ for the remaining 15~\%.

In Table~\ref{table:props_dsf} we present the results of density for the strategies mentioned at the beginning of this section.
When comparing cases (1) and (2) for each model, we note that the density values practically do not differ.
In addition, the model of Koller \textit{et al.} \cite{Koller_2012} best reproduces the experimental value.
The model due to Weber and Kirchner also presents a small deviation of 1\%.
We have also computed relevant radial distribution functions (RDFs), which can be found in the Supplementary material.
The better agreement is found for the models of Koller \textit{et al.} \cite{Koller_2012} and Batista \textit{et al.} \cite{Batista_2015}.
Despite the small deviations observed for the model of Liu \textit{et al.} \cite{Liu_2014}, the simulation with rigid bodies qualitatively reproduces the RDFs obtained for the flexible ions.
Although not shown here, these conclusions are valid for other calculated RDFs: N(anion)-N(cation), C6(cation)-N(anion), C2(cation)-B(anion).
The model of Weber and Kirchner \cite{Weber_2016}, on the other hand, presents important discrepancies in the RDFs of the pairs C5(cation)-N(anion) and C2(cation)-B(anion) when using rigid bodies.
That is why, in this case only the imidazolium ring was considered rigid and the ethyl chain was left unconstrained.

@@

A study of dihedrals was also carried out for \ce{[emim][NTf_2]}.
The CDF of the dihedrals angles of C1-N2-C5-C6 and C2-N2-C5-C6, shown in Fig.~\ref{fig:die_ntf2}(a), resembles the results of Fig.~\ref{fig:dihedrals_emim}(d), but the most likely angle of C1-N2-C5-C6 is 0$^{\circ}$.
This similarity was expected as both models of \ce{[emim]^+} are based on the force field developed by Canongia-Lopes and P\'{a}dua~ \cite{Canongia_Lopes_2006}.
With regards to the anion \ce{[NTf_2]^-}, Deetlefs \textit{et al.} \cite{Deetlefs_2006} have experimentally determined the conformer population of the liquid phase and observed that \ce{[NTf_2]^-} is mostly in its \textit{trans} conformation.
Fig.~\ref{fig:die_ntf2}(b), which corresponds to the CDF for the contiguous C1-S1-N1-S2 and S1-N1-S2-C2 chains, qualitatively verifies the observation of Deetlefs \textit{et al.} \cite{Deetlefs_2006}.
Given the considerable flexibility of \ce{[NTf_2]^-}, a more rigorous strategy is required for building the rigid bodies.
Thus, the simulations of \ce{[emim][NTf_2]} were performed considering flexible ions and the DSF method, which results in a computed density of 1.495 $\pm$ 0.009 g/cm$^3$ with an error of 1.12\% with respect to the experimental value~\cite{Tokuda_2005}.

@@

In Table \ref{table:henry} we present the results of solvation free energies of \ce{CO_2} along with the Henry constants ($K_s$) obtained by means of Eq.~\eqref{eq:henry_eq}.
Table \ref{table:henry} also includes the results for [emim][Tf\textsubscript{2}N] as well as the experimental $K_s$ reported by Mahurin \textit{et al.} \cite{Mahurin_2010} and Finotello \textit{et al.} \cite{Finotello_2008}.
The uncertainties in $K_s$ were calculated using standard error propagation.

@@

From Table \ref{table:henry} one can see that the model of Koller \textit{et al.} \cite{Koller_2012} best reproduces the experimental value of $K_s$.
If we look at the $\Delta G_{\lambda_{\,\text{vdW}}}$ profiles in Fig.~\ref{fig:deltag}, we observe that for the models of Batista \textit{et al.} \cite{Batista_2015} and Liu \textit{et al.}~\cite{Liu_2014} the maximum occurs at lower values of energy, suggesting that these models underestimate the interatomic interactions, which may also explains their failure to reproduce the experimental densities.

@@

The $\Delta G_{\, \text{sim}}$ values for the models of Koller \textit{et al.} \cite{Koller_2012} and K\"{o}ddermann \textit{et al.} \cite{K_ddermann_2007} suggest a more favorable solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf\textsubscript{2}]}. 
The spatial distribution functions (SDFs) \cite{Svishchev_1993} shown in Figs.~\ref{fig:sdf_ions} confirm these results.
The SDFs of the ions around a \ce{CO_2} molecule were obtained by using the software packages TRAVIS \cite{Brehm_2011} and VMD \cite{HUMP96}.
For that, we carried out NPT-MD simulations of the systems \ce{CO_2}-\ce{[emim][B(CN)_4]} and \ce{CO_2}-[emim][Tf\textsubscript{2}N] with a solute molar fraction of 0.1.
The isovalues were visually determined so as to observe the first solvation shell.
From Fig.~\ref{fig:sdf_ions}(b), which correspond to the ions \ce{[emim]^+} and \ce{[NTf_2]^-}, it can be seen that the region with a low probability of finding an ion is larger than for \ce{[emim][B(CN)_4]}.
Also, the first solvation shell due to [B(CN)\textsubscript{4}]$^{-}$ occurs at a sligthly shorter distance than for \ce{[NTf_2]^-}, as one can note from Fig.
\begin{figure}[ht]
\centering
\subfloat[]{%
  \includegraphics[width=\linewidth]{kollerall.pdf}%
}

@@

Having selected the model of \ce{[emim][B(CN)_4]} that best reproduces the experimental $K_s$ of \ce{CO_2}, we carried out the calculation of infinite-dilution activity coefficients ($\gamma^{\, \infty}$) of benzene, hexane, cyclohexane, ethanol and water in \ce{[emim][B(CN)_4]}.
These solutes were modeled as described in Sec.~\ref{sec:force_field_sol}.

@@

In Secs.~\ref{sec:henry_results} and \ref{sec:act_results} we obtained some solvation properties for diluted solutions. However, the corresponding derived conclusions cannot be extraploted to concentrated solutions, which are the more likely scenario in the \ce{CO_2} capture. Therefore, in this section we proceed to analyze the \ce{CO_2} solvation at the thermodynamic conditions specified in Table \ref{table:solv}. These conditions correspond to close values of molality of \ce{CO_2} in \ce{[emim][B(CN)_4]} and [emim][Tf\textsubscript{2}N], which is the solubility measure employed by Carvalho \textit{et al.}~\cite{Carvalho_2016}.

@@

In Fig. we present the results of each iteration according to the scheme proposed in Sec. \ref{sec:partial_pressure}. The predicted values of pressure ($P_s$) reproduce what is observed experimentaly, in the sense that $P_s$ for the system \ce{CO_2}-\ce{[emim][B(CN)_4]} is considerably smaller than the value obtained for \ce{CO_2}-[emim][Tf\textsubscript{2}N].% However, in this case we were not able to converge because an oscilatory behaviour was observed.

\begin{table*}
\centering
\begin{adjustbox}{width=1\textwidth}
\begin{threeparttable}
\caption{Gaseous phase pressure, predicted from solvation free energies of \ce{CO_2} for the systems \ce{CO_2}-\ce{[emim][B(CN)_4]} and \ce{CO_2}-[emim][Tf\textsubscript{2}N].}
\begin{tabular}{ c c  c  c  c  c c }  
\toprule
\# & P &$\Delta G_{\,\text{LJ}}$  & $\Delta G_{\,\text{Coul}}$  & $\Delta G_{\,\text{sim}}$ & $f^{g}_s$ (Eq.~\eqref{eq:fgas_d}) & $P_s$ (Ec.~\eqref{eq:fgas}) \\
 &(atm) & (kcal/mol) & (kcal/mol) &  (kcal/mol) & (atm) & (atm) \\
		\hline
		\multicolumn{7}{c}{\ce{[emim][B(CN)_4]}~\cite{Koller_2012} } \\
1 & 21.0 &  0.42 $\pm$ 0.01 & -1.176  $\pm$ 0.007  & -0.755 $\pm$ 0.01 & 16.96 $\pm$ 0.33 & 18.33 $\pm$ 0.4 \\
2 & 18.33 &  0.39 $\pm$ 0.01 & -1.187  $\pm$ 0.008  & -0.794 $\pm$ 0.01 & 15.93 $\pm$ 0.33 & 17.14 $\pm$ 0.4 \\ 
3 & 17.14 &  0.39 $\pm$ 0.01 & -1.164  $\pm$ 0.008  & -0.779 $\pm$ 0.01 & 16.32 $\pm$ 0.33 & 17.5 $\pm$ 0.4 \\
		\hline
       \multicolumn{7}{c}{[emim][Tf\textsubscript{2}N]~\cite{K_ddermann_2007} } \\
1 & 37.0 &  0.314 $\pm$ 0.005  & -0.856 $\pm$ 0.005 & -0.542 $\pm$ 0.005 & 32.6 $\pm$ 0.37 & 38.4 $\pm$ 0.4 \\
2 & 38.4 &  0.349 $\pm$ 0.005  & -0.840  $\pm$ 0.005 & -0.491 $\pm$ 0.005 & 35.35  & 42.47 \\
 \bottomrule
\label{table:deltaG_conc} 
\end{tabular}
\end{threeparttable}
\end{adjustbox}
\end{table*}

>>>>

The initial configurations were obtained with PLAYMOL \cite{playmol}, a software integrated with PACKMOL \cite{Mart_nez_2009}, using the tool \textit{build} to specify geometric data (bond and angle lengths, and dihedral angles), from which PLAYMOL automatically determines the Cartesian coordinates.
We first carried out NVT MD simulations of the isolated ion pairs in order to obtain a minimum energy conformation.
For this, we used the Nos\'{e}-Hoover thermostat \cite{Martyna1992} and a cubic box of length L = 50 {\AA} with non-periodic boundary conditions.
After a short run with $\Delta t$ = $0.5~\text{fs}$, we carried out a ``quenching'' simulation with $\Delta t$ = $1 ~\text{fs}$, in which the temperature is gradually reduced until reaching a value close to $0~\text{K}$.
The resulting coordinates were then specified in PLAYMOL \cite{playmol} to build the configurations of the liquid phases.

@@

Using the software TRAVIS \cite{Brehm_2011}, we obtained the combined distribution functions (CDFs) of the dihedral angles of C1-N2-C5-C6 and C2-N2-C5-C6, as sampled in NPT-MD simulations of 250 ion pairs at $T = 298.0~\text{K}$ and $P = 1.0~\text{bar}$.
In Fig.~\ref{fig:dihedrals_emim} we show the CDFs for each force field of \ce{[emim][B(CN)_4]}.
A first inspection of this figure reveals a similar behavior of the models due to Koller \textit{et al.} \cite{Koller_2012} and Liu \textit{et al.} \cite{Liu_2014}.
This was expected as both are based on the AMBER framework \cite{Cornell_1995}.
In addition, Figs.~\ref{fig:dihedrals_emim}(a) and \ref{fig:dihedrals_emim}(b) are symmetric, which means that atoms C1, N2, and C2 are coplanar.
In Fig.~\ref{fig:dihedrals_emim}(a), the most likely angles can be found in a narrow region around 90$^{\circ}$ and 270$^{\circ}$.
Since these angles correspond to an equivalent configuration in the \textit{united-atom} \ce{[emim]^+}, in the case of Koller \textit{et al.} \cite{Koller_2012} we built all the rigid bodies by specifying 90$^{\circ}$ for the dihedral angle of C6-C5-N2-C1.
The other models show a larger flexibility with high probability regions that extend in a broad range of angle values.
Despite this, it was possible to build more than one type of rigid body by specifying the most probables angles of C6-C5-N2-C1.
In the case of Liu \textit{et al.} \cite{Liu_2014}, these angles correspond to 90$^{\circ}$ and -90$^{\circ}$, while for the model of Batista \textit{et al.} \cite{Batista_2015}, we employed 102.6$^{\circ}$ and -102.6$^{\circ}$.
These two types of rigid bodies were built in equal proportion.
Fig.~\ref{fig:dihedrals_emim}(d) corresponds to Weber and Kirchner \cite{Weber_2016}. 
Given the wide range of sampled angles, we defined four types of rigid bodies: 150$^{\circ}$, 180$^{\circ}$ and 210$^{\circ}$ were employed to build 30~\%, 25~\% and 30~\% of them, respectively; and 0$^{\circ}$ for the remaining 15~\%.

In Table~\ref{table:props_dsf} we present the results of density for the strategies mentioned at the beginning of this section.
When comparing cases (1) and (2) for each model, we note that the simulated densities practically do not differ.
In addition, the models of Koller \textit{et al.} \cite{Koller_2012} and Weber and Kirchner \cite{Weber_2016} best reproduce the experimental value.
We have also computed radial distribution functions (RDFs), which can be found in the Supplementary material.
The better agreement is found for the models of Koller \textit{et al.} \cite{Koller_2012} and Batista \textit{et al.} \cite{Batista_2015}.
Despite the small deviations observed for the model of Liu \textit{et al.} \cite{Liu_2014}, the simulation with rigid bodies qualitatively reproduces the RDFs obtained for the flexible ions.
Although not shown here, these conclusions are valid for other calculated RDFs: N(anion)-N(cation), C6(cation)-N(anion), C2(cation)-B(anion).
The model of Weber and Kirchner \cite{Weber_2016}, on the other hand, presents important discrepancies in the RDFs of the pairs C5(cation)-N(anion) and C2(cation)-B(anion) when using rigid bodies.
That is why, in this case only the imidazolium ring was considered rigid and the ethyl chain was left unconstrained.

@@

A study of dihedrals was also carried out for \ce{[emim][NTf_2]}.
The CDF of the dihedrals angles of C1-N2-C5-C6 and C2-N2-C5-C6, shown in Fig.~\ref{fig:die_ntf2}(a), resembles the results of Fig.~\ref{fig:dihedrals_emim}(d), but the most likely angle of C1-N2-C5-C6 is 0$^{\circ}$.
This similarity was expected as both models of \ce{[emim]^+} are based on the force field developed by Canongia-Lopes and P\'{a}dua~ \cite{Canongia_Lopes_2006}.
With regards to the anion \ce{[NTf_2]^-}, Deetlefs \textit{et al.} \cite{Deetlefs_2006} have experimentally determined the conformer population of the liquid phase and observed that \ce{[NTf_2]^-} is mostly in its \textit{trans} conformation.
Fig.~\ref{fig:die_ntf2}(b), which corresponds to the CDF for the contiguous C1-S1-N1-S2 and S1-N1-S2-C2 chains, qualitatively verifies the observation of Deetlefs \textit{et al.} \cite{Deetlefs_2006}.
Given the considerable flexibility of \ce{[NTf_2]^-}, a more rigorous strategy for building the rigid bodies is required.
Thus, the simulations of \ce{[emim][NTf_2]} were performed considering flexible ions and the DSF method, which results in a computed density of 1.495 $\pm$ 0.009 g/cm$^3$ with an error of 1.12\% with respect to the experimental value~\cite{Tokuda_2005}.

@@

In Table \ref{table:henry} we present the results of solvation free energies of \ce{CO_2} along with the Henry constants ($K_s$) obtained by means of Eq.~\eqref{eq:henry_eq}.
Table~\ref{table:henry} also contains the experimental $K_s$ reported by Mahurin \textit{et al.} \cite{Mahurin_2010} and Finotello \textit{et al.} \cite{Finotello_2008}.
The uncertainties in $K_s$ were calculated using standard error propagation.

@@

From Table \ref{table:henry} one can see that the model of Koller \textit{et al.} \cite{Koller_2012} best reproduces the experimental value of $K_s$.
If we look at the $\Delta G_{\lambda_{\,\text{vdW}}}$ profiles in Fig.~\ref{fig:deltag}, we observe that for the models of Batista \textit{et al.} \cite{Batista_2015} and Liu \textit{et al.}~\cite{Liu_2014} the maximum occurs at lower values of energy, suggesting that these models underestimate the interatomic interactions, which may also explain their failure to reproduce the experimental density.

@@

The $\Delta G_{\, \text{sim}}$ values for the models of Koller \textit{et al.} \cite{Koller_2012} and K\"{o}ddermann \textit{et al.} \cite{K_ddermann_2007} suggest a more favorable solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} than in \ce{[emim][NTf\textsubscript{2}]}. 
The spatial distribution functions (SDFs) \cite{Svishchev_1993} shown in Figs.~\ref{fig:sdf_ions} confirm this result.
The SDFs of the ions around a \ce{CO_2} molecule were obtained by using the software packages TRAVIS \cite{Brehm_2011} and VMD \cite{HUMP96}.
For that, we carried out NPT-MD simulations of \ce{CO_2}-\ce{[emim][B(CN)_4]} and \ce{CO_2}-[emim][Tf\textsubscript{2}N] systems with a \ce{CO_2} molar fraction of 0.1.
The isovalues were visually determined so as to observe the first solvation shell.
If we compare Figs.~\ref{fig:sdf_ions} (a) and (b), one notes that the region with a low probability of finding an ion is larger for \ce{[emim][NTf_2]}.
Also, the first solvation shell due to [B(CN)\textsubscript{4}]$^{-}$ occurs at a slightly shorter distance than for \ce{[NTf_2]^-}, as one sees from Fig.
\begin{figure}[ht]
\centering
\subfloat[]{%
  \includegraphics[width=\linewidth]{kollerall.pdf}%
}

@@

Having selected the model of \ce{[emim][B(CN)_4]} that best reproduces the experimental $K_s$ of \ce{CO_2}, we next calculate the infinite-dilution activity coefficients ($\gamma^{\, \infty}$) of benzene, hexane, cyclohexane, ethanol and water in \ce{[emim][B(CN)_4]}.

@@

Since concentrated solutions are the more likely scenario in \ce{CO_2} capture processes, we carried out simulations at the thermodynamic conditions specified in Table \ref{table:solv}, which correspond to close values of \ce{CO_2} molality in \ce{[emim][B(CN)_4]} and [emim][Tf\textsubscript{2}N].

@@

In Fig. we present the results of each iteration according to the scheme proposed in Sec. \ref{sec:partial_pressure}. The predicted values of pressure ($P_s$) reproduce what is observed experimentaly, in the sense that $P_s$ for the system \ce{CO_2}-\ce{[emim][B(CN)_4]} is considerably smaller than the value obtained for \ce{CO_2}-[emim][Tf\textsubscript{2}N].

\begin{figure}[ht]
\includegraphics[width=\linewidth]{pressure_est.pdf}
\caption{$\Delta G_{\lambda_{\,\text{vdW}}}$ profiles corresponding to the insertion of a \ce{CO_2} molecule in \ce{[emim][B(CN)_4]}, which was modeled according to the force fields of (circles) Koller \textit{et al.} \cite{Koller_2012}, (squares) Batista \textit{et al.} \cite{Batista_2015}, (up triangles) Liu \textit{et al.} \cite{Liu_2014}, and (x) Weber y Kirchner \cite{Weber_2016}.}
\label{fig:pressure}
\end{figure}
